NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE, PG and PS lipids

### 2017-11-05 Manuscript/manuscript.tex
Update figures
Add lipids figure
Add dihedral plots by Buslaev into manuscript
Fix error in data due to mistake in taking data from scherer87.

@@

\subsection{Effect of PE, PS and PG on PC headgroup}
The headgroup order parameters for PC lipids (POPC and DOPC)
mixed with PE, PS and PG lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
from different simulation model and experiments \cite{scherer87} with different
mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
This is not the case in simulation data shown in Fig. \ref{HGorderparametersPCvsPEPSPG}.
The addition of DOPE into a POPC and DOPC bilayers significantly decreases the PC headgroup
order parameters in simulations with OPLS compatible version of the Berger force field \cite{tieleman06}
in contrast to experiments \cite{scherer87}. On the other hand, the increase of the PC
headgroup order parameters in CHARMM36 simulations mixed with PS and PG lipids is significantly
smaller than in experiments.
\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPSPG.eps}
  \caption{\label{HGorderparametersPCvsPEPSPG}
    PC headgroup order parameters from mixtures with PE, PS and PG
    lipids with various mole fractions from different simulation models and experiments \cite{scherer87}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure*}

@@

Comparison of Ca$2+$ binding in PG between CHARMM36 simulations and experiments \cite{borle85}
is shown in Fig. \ref{changesWITHCaClPG}. The decrease of $\alpha$ order parameter
is in agreement with experiments, while decerase of $\beta$ order parameter is overestimated.
The result is very similar to the results with PC in NMRlipids II publication \cite{catte16}.
It should be, however, noted that the $\beta$-order parameters are not actually measured for PG,
but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
parameter of bound charge \todo{Should we check this against cationic surfactant experiments \cite{scherer89}?}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CHANGESwithCaClPG.eps}
  \caption{\label{changesWITHCaClPG}
    PG order parameters as a function CaCl$_2$ concentration from experiments \cite{borle85} and CHARMM36 simulations.
    Note that beta order parameter is calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}, not actually measured. 
  }
\end{figure}

@@

>>>>

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
  \caption{\label{lipids}
    Chemical structures and labels for the headgroup carbons.
  }
\end{figure}

@@

\subsection{Effect of PE, PS and PG on PC headgroup}
The headgroup order parameters for PC lipids (POPC and DOPC)
mixed with PE, PS and PG lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
from different simulation model and experiments \cite{scherer87} with different
mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
This is not the case in simulation data shown in Fig. \ref{HGorderparametersPCvsPEPSPG}.
The addition of DOPE into a POPC and DOPC bilayers significantly decreases the PC headgroup
order parameters in simulations with OPLS compatible version of the Berger force field \cite{tieleman06}
in contrast to experiments \cite{scherer87}. On the other hand, the increase of the PC
headgroup order parameters in CHARMM36 simulations mixed with PS and PG lipids is significantly
smaller than in experiments.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPCvsPEPSPGchol.eps}
  \caption{\label{HGorderparametersPCvsPEPSPGchol}
    PC headgroup order parameters from experiments of mixtures with
    PE, PS, PG and cholesterol \cite{scherer87,scherer89,ferreira13}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
\end{figure}

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPSPG.eps}
  \caption{\label{HGorderparametersPCvsPEPSPG}
    PC headgroup order parameters from mixtures with PE, PS and PG
    lipids with various mole fractions from different simulation models and experiments \cite{scherer87}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure*}

@@

Comparison of Ca$2+$ binding in PG between CHARMM36 simulations and experiments \cite{borle85}
is shown in Fig. \ref{changesWITHCaClPG}. The decrease of $\alpha$ order parameter
is in agreement with experiments, while decerase of $\beta$ order parameter is overestimated.
The result is very similar to the results with PC in NMRlipids II publication \cite{catte16}.
It should be, however, noted that the $\beta$-order parameters are not actually measured for PG,
but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
parameter of bound charge \todo{Should we check this against cationic surfactant experiments \cite{scherer89}?}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CHANGESwithCaClPGPS.eps}
  \caption{\label{changesWITHCaClPG}
    PG order parameters as a function CaCl$_2$ concentration from experiments \cite{borle85} and CHARMM36 simulations.
    Note that beta order parameter is calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}, not actually measured. 
  }
\end{figure}

@@

\section{Dihedrals}
\begin{figure*}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/dihed1.png}
  \includegraphics[width=8.0cm]{../Figs/dihed2.png}
  \includegraphics[width=8.0cm]{../Figs/dihed3.png}
  \includegraphics[width=8.0cm]{../Figs/dihed4.png}
  \includegraphics[width=8.0cm]{../Figs/dihed5.png}
  \includegraphics[width=8.0cm]{../Figs/dihed6.png}
  \includegraphics[width=8.0cm]{../Figs/dihed7.png}
  \includegraphics[width=8.0cm]{../Figs/dihed8.png}
  \includegraphics[width=8.0cm]{../Figs/dihed9.png}
  \includegraphics[width=8.0cm]{../Figs/dihed10.png}
  \caption{\label{dihedrals}
    Experimental results for sign measurement for POPS sample
  }
\end{figure*}
